---
title: "Avian Interaction Figure Data Processing"
format: 
  html: 
    self-contained: true
editor: visual
---

This script provides centralized data loading, cleaning and utility functions that are used in the phylogeny, network and interaction distribution figures. The pre-processing needed for each figure is walked through, in the order they appear in the paper. For actual plotting, see `summary_vignette`

First, we load the packages and datasets required for all three figures

```{r}
#| warning: false

rm(list=ls())
# library(tidyverse)
library(tidyr)
library(dplyr)
library(stringr)
library(clootl)
library(ggtree)
library(ggplot2)
library(ggnewscale)
library(igraph)
library(ggraph)
library(cowplot)
library(jsonlite)
library(auk)
library(magick)

inter_NA <- read.csv(here::here("../AvianMetaNetwork-Working/L1/amn_cac.csv")) #NA avian interaction data
```

The distribution figure (Fig. 2) and example networks (Fig. 4) both employ a mutual color scheme and order of interactions. This is defined in the `interaction_categories_and_colors.R` script and called here:

```{r}
source(here::here("./R/lib/interaction_categories_and_colors.R"))
```

One thing to note about the database is that it includes many rows with duplicate interactions (same species pair and interaction type). For example:

```{r}
print(inter_NA[1:2,1:10])
```

These are functionally the same, and only differ by their text excerpts!

This was done such that each piece of evidence is reflected, but causes issues with determining the absolute amount of interactions present. Because of this, we will created a second version of `inter_NA` that removes interactions with the same species pair and interaction type. This will be tagged `inter_NA_slim`

```{r}
inter_NA_slim <- inter_NA %>%
  rowwise() %>%
  mutate(
    #Create columns of the species pairs in alphabetical order
    sp_min = min(taxa1_clements, taxa2_clements),
    sp_max = max(taxa1_clements, taxa2_clements),
    #Check if order is swapped
    swapped = taxa1_clements > taxa2_clements,
    #Normalize effects based on alphabetical order
    eff_min_on_max = if_else(swapped, effect_tx2_on_tx1, effect_tx1_on_tx2),
    eff_max_on_min = if_else(swapped, effect_tx1_on_tx2, effect_tx2_on_tx1)
  ) %>%
  ungroup() %>%
  #Remove duplicates based on normalized columns
  distinct(sp_min, sp_max, interaction, eff_min_on_max, eff_max_on_min, 
           .keep_all = TRUE) %>%
  #Delete the helper columns
  select(-sp_min, -sp_max, -swapped, -eff_min_on_max, -eff_max_on_min)

message("We removed ", nrow(inter_NA) - nrow(inter_NA_slim), " duplicate rows!")
```

## Distribution of Interactions by Type (Figure 2)

The distribution figure (fig. 2) summarizes the whole, raw dataset. The only further step is generating a summary dataframe that counts the number of instances of each interaction type. We will do this for both the slim and raw datasets:

```{r}
type_summ <- inter_NA %>%
  group_by(interaction) %>%
  summarize(total = n()) %>%
  left_join(interaction_categories, by = "interaction") %>%
  arrange(category, desc(total))

#Repeat for slim dataset
type_summ_slim <- inter_NA_slim %>%
  group_by(interaction) %>%
  summarize(unique = n()) %>%
  left_join(interaction_categories, by = "interaction") %>%
  arrange(category, desc(unique))


#Join to have total and unique for each type
type_summ_total <- merge(type_summ, type_summ_slim, sort = FALSE)
type_summ_total$interaction <- factor(type_summ_total$interaction,
                                     levels = rev(type_summ_total$interaction))

head(type_summ_total)
```

Now, we can plot the distribution figure!

## Phylogeny Figure (Figure 3)

For both the phylogeny figure (Fig. 3) and network figure (Fig. 4), we do some further processing. Firstly, we remove non-specific taxa (unidentified species and hybrids) which won't map to tips on the phylogenetic tree. The same can be said for subspecies, so we generalize subspecies back to their species level identification. This causes some "self interactions" where a species reacts with its subspecies, so those are removed as well.

```{r}
#load in clements data for identifying subspecies and species
spp <- read.csv(here::here("./data/L1/species_checklists/spp_clem_in_amn_cac.csv")) 

inter_NA_trim <- inter_NA_slim %>%
    #Join with spp to get category information for both taxa
    left_join(spp %>% select(scientific.name, category), 
              by = c("taxa1_clements" = "scientific.name")) %>%
    rename(taxa1_category = category) %>%
    left_join(spp %>% select(scientific.name, category), 
              by = c("taxa2_clements" = "scientific.name")) %>%
    rename(taxa2_category = category) %>%
    #Keep rows where both taxa1 and taxa2 are subspecies
    filter(
      taxa1_category %in% c("species", "subspecies"),
      taxa2_category %in% c("species", "subspecies")
    ) %>%
    #Collapse all to species level (simplifying subspecies)
    mutate(
      taxa1_clements = word(taxa1_clements, 1, 2),
      taxa2_clements = word(taxa2_clements, 1, 2)
    ) %>%
    #Remove the temporary category columns
    select(-taxa1_category, -taxa2_category) %>%
    #Remove self interactions
    filter(taxa1_clements != taxa2_clements)

message("After initial cleaning, the data contains ", nrow(inter_NA_trim), " rows")
```

We will also repeat the slimming step, as collapsing the subspecies to their larger species creates new duplicate rows (same species pair and interaction type). For example, `Acanthis flammea hornemanni/exilipes` is converted to `Acanthis flammea`, which may cause new duplicated interactions, where a species had separate interactions with both the species and subspecies.

```{r}
inter_NA_clean <- inter_NA_trim %>% 
  rowwise() %>% #This removes duplicate interactions (same sp pair and int type)
  mutate(       #by creating columns of the species pairs in alphabetical order
    sp_min = min(taxa1_clements, taxa2_clements),
    sp_max = max(taxa1_clements, taxa2_clements)
  ) %>%
  ungroup() %>%
  distinct(sp_min, sp_max, interaction, .keep_all = TRUE) %>% #removing duplicates
  select(-sp_min, -sp_max) #and deleting the helper columns

message("We removed ", nrow(inter_NA_trim) - nrow(inter_NA_clean), " new duplicates caused by subspecies collapse")
```

Currently, our data is formatted like this:

|       |              |       |
|-------|:-------------|------:|
| taxa1 | interaction1 | taxa2 |
| taxa1 | interaction2 | taxa2 |

We want to combine it into two columns, for easier summary of total interactions per species in the phylogeny, like so:

|          |              |
|----------|:-------------|
| taxa1    | interaction1 |
| taxa2    | interaction1 |
| taxa1    | interaction2 |
| species3 | interaction2 |

This duplicates our dataframe, so that every interaction: sp1 int1 sp2 is expressed in both sp1/sp2 permutations.

```{r}
#| warning: false

inter_NA_flip <- inter_NA_clean %>%
  rename(
      temp_sp1_com = .data[["taxa2_common"]],
      temp_sp2_com = .data[["taxa1_common"]],
      temp_sp1_cle = .data[["taxa2_clements"]],
      temp_sp2_cle = .data[["taxa1_clements"]]
    ) %>%
    rename(
      taxa1_common = temp_sp1_com,
      taxa2_common = temp_sp2_com,
      taxa1_clements = temp_sp1_cle,
      taxa2_clements = temp_sp2_cle
    )

inter_NA_full <- dplyr::union(inter_NA_clean, inter_NA_flip)

message("Now our database has increased to ", nrow(inter_NA_full), " rows, with duplicates automatically removed")
```

Union will automatically discard identical rows, and our initial cleanings ensure there are no species pair/interaction type duplicates left in the dataframe!

Next, we simplify the dataframe for the phylogeny visualization. We create a new data frame with one row per species, column "n_int" for the total number of unique interactions that that species participates in, and "n_type" for the total number of distinct types of interactions per species.

```{r}
inter_NA_working <- inter_NA_full %>%
  group_by(.data[["taxa1_clements"]]) %>%
  #Replace spaces with underscores to match tree tips
  mutate(taxa1_scientific = str_replace_all(.data[["taxa1_clements"]], " ", "_")) %>%
  rename(species = taxa1_clements) %>%
  summarize(
    n_int = n(),
    n_type = n_distinct(.data[["interaction"]]),
    .groups = "drop"
  )

head(inter_NA_working)
```

Finally, we will attach each species a family from the Clements Checklist 2024.

```{r}
inter_NA_working <- inter_NA_working %>%
  left_join(
    spp %>% 
      mutate(genus_species = word(scientific.name, 1, 2)) %>%
      select(genus_species, family), 
    by = c("species" = "genus_species")
  )

inter_NA_working <- inter_NA_working %>%
  mutate(species = str_replace_all(species, " ", "_"))
```

Now we can use `clootl` to extract a phylogenetic tree. We need to remove the underscores from the species names temporarily to search with `clootl`, but these are left in the `inter_NA_working` dataframe to match with the tip labels later.

```{r}
inter_NA_list <- str_replace_all(inter_NA_working$species, "_", " ")
NA_tree <- extractTree(species = inter_NA_list, label_type = "scientific", taxonomy_year = 2024)
```

We now have everything we need to plot the phylogeny! We use the following function to plot.

```{r}
plot_phylo_combined <- function(tree, #tree produced earlier, should be NA_tree
                                data, #working dataset
                                value_col = "n_int", #value column, can replace with "n_type"
                                label_offset = 8, #offset of family labels
                                label_size = 4.5, #size of family labels
                                tip_size = 2, #size of colored circles representing families
                                min_species = NULL, #optional species size threshold for labeling
                                legend_title = "") { #legend title

  #Rename species column to match requirements in ggtree
  plot_data <- data %>%
    rename(label = .data[["species"]])

  #Calculate family size for filtering
  family_metrics <- plot_data %>%
    filter(!is.na(.data[["family"]])) %>%
    group_by(.data[["family"]]) %>%
    summarise(
      n_species = n(),
      .groups = "drop"
    )

  #Determine which families to label based on species threshold
  families_to_label <- family_metrics

  if (!is.null(min_species)) {
    families_to_label <- families_to_label %>%
      filter(n_species >= min_species)
  }

  families_to_label <- families_to_label %>%
    pull(.data[["family"]])

  #Use tree tip order to find true phylogenetic center
  tree_tips <- tree$tip.label

  #Calculate family positions
  family_positions <- plot_data %>%
    filter(!is.na(.data[["family"]]),
           .data[["family"]] %in% families_to_label) %>%
    mutate(tree_order = match(label, tree_tips)) %>%
    group_by(.data[["family"]]) %>%
    summarise(
      median_pos = median(tree_order, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    arrange(median_pos)

  #Get the central (median position) species per labeled family for labeling
  rep_df <- plot_data %>%
    filter(!is.na(.data[["family"]]),
           .data[["family"]] %in% families_to_label) %>%
    mutate(tree_order = match(label, tree_tips)) %>%
    left_join(family_positions, by = "family") %>%
    group_by(.data[["family"]]) %>%
    arrange(abs(tree_order - median_pos)) %>%
    slice(1) %>%
    ungroup() %>%
    select(label, family = .data[["family"]])


  #Get family order
  family_order <- family_positions %>%
    pull(.data[["family"]])

  n_labeled <- length(family_order)

  #Build a custom palette
  all_colors <- c(
    "#FF7F00", "#111222", "#7570B3", "#E7298A", "#66A61E", "#E6AB02",
    "#A6761D", "#666666", "#E41A1C", "#377EB8", "#4DAF4A", "#984EA3",
    "#1B9E77", "#A65628", "#F781BF", "#999999", "#8DD3C7", "#FB8072",
    "#80B1D3", "#FDB462", "#BC80BD"
  )

  #Assign colors sequentially in phylogenetic order
  if (n_labeled <= length(all_colors)) {
    family_palette <- all_colors[1:n_labeled]
  } else {
    #If more families than colors, repeat the palette
    family_palette <- rep(all_colors, length.out = n_labeled)
  }

  #Create named vector with colors assigned by phylogenetic order
  family_colors <- setNames(family_palette, family_order)

  #Filter plot_data to only include labeled families for tip coloring
  plot_data_tips <- plot_data %>%
    mutate(family_filtered = ifelse(.data[["family"]] %in% families_to_label,
                                    .data[["family"]],
                                    NA_character_))

  #Set breaks for the legend
  val_range <- range(plot_data[[value_col]], na.rm = TRUE)
  breaks <- c(val_range[1], 5, 25, 100, val_range[2])

  #Create base tree plot
  p <- ggtree(tree, layout = "circular") %<+% plot_data +
    geom_tree(aes(color = .data[[value_col]])) +
    scale_color_gradientn(
      trans = "log10",
      colors = c("#440154FF", "#414487FF", "#2A788EFF", "#22A884FF", "#7AD151FF", "#FDE725FF"),
      breaks = breaks,
      labels = breaks,
      name = legend_title
    ) +
    ggnewscale::new_scale_color() +
    #Only plot tips for families that meet threshold
    geom_tippoint(data = . %>% filter(.data[["family"]] %in% families_to_label),
                  aes(color = .data[["family"]]),
                  size = tip_size) +
    scale_color_manual(values = family_colors, na.value = "grey50", name = "Family", guide = "none") +
    theme(
      legend.position = c(0.735, 0.26),
      legend.justification = c(1, 0),
      legend.title = element_text(size = 20),
      legend.text = element_text(size = 18)
    )

  # Add family labels at center positions
  p <- p +
    ggnewscale::new_scale_color() +
    geom_tiplab2(
      data = . %>% filter(label %in% rep_df$label),
      aes(label = .data[["family"]], color = .data[["family"]]),
      size = label_size,
      offset = label_offset,
      hjust = 0,
      angle = 0,
      fontface = "bold"
    ) +
    scale_color_manual(values = family_colors, na.value = "grey50", guide = "none")

  return(p)
}
```

## Interaction Network Figure (Fig. 4)

The interaction figures use creative commons images sourced from AviCommons, using the following files:  `aux_avicommons_full.json` and `clemtax_2025.csv`. These need to be sourced into the environment and we need to define a couple of functions to create the images:

```{r, warning=F, message=F}

source(here::here("R/lib/avicommons_clem.r"))
avi <- jsonlite::fromJSON(here::here("data/L2/aux_avicommons_full.json"))
clem <- readr::read_csv(here::here("data/L1/species_checklists/spp_clem_in_amn_cac.csv"))

```

Next, we need edit `inter_NA_clean` to remove a few interaction types (co-occurrence and hybridization) that don't fit well into an ecological network.

```{r}
inter_NA_int <- inter_NA_clean %>% #removing "non-interactions"
  filter(interaction != "co-occur" & interaction != "hybridization")
```

Now we can plot! We provide two functions. the first plots a network for a single species, showing all of it's interacting species and the interactions between them.

```{r}
create_network <- function(data, #input dataset, should be inter_NA_int
                           focal_species, #scientific name of a species of interest
                           layout = "fr", #geometric layout algorithm for plot
                           show_labels = TRUE, #option to label species names
                           image_size = 480, #size of bird image, options are 240, 320, 480, 900
                           bird_width = 0.3, #relative size of bird image
                           network_width = 0.7, #relative size of network
                           edge_alpha = 0.5, #alpha value of interactions (transparency)
                           edge_width = 1.25, #linewidth of interactions
                           focal_node_size = 10, #size of species of interest
                           node_size = 8, #size of interacting species
                           node_stroke = 1.2, #node outline width
                           focal_node_stroke = 1.5, #focal species outline width
                           focal_label_size = 5, #focal species text size (if show_labels = T)
                           label_size = 4, #interacting species text size (if show_labels = T)
                           legend_text_size = 23, #interaction type text size
                           legend_title_size = 25, #legend title text size
                           curve_edges = FALSE, #set true for visualizing multiple int. between two species
                           cats = FALSE) { #set true to color by category instead of interaction type

  #Filter interactions involving focal species
  focal_interactions <- data %>%
    filter(taxa1_clements == focal_species |
             taxa2_clements == focal_species)

  #Get network species and their interactions
  network_species <- unique(c(focal_interactions$taxa1_clements,
                              focal_interactions$taxa2_clements))

  network_data <- data %>%
    filter(taxa1_clements %in% network_species,
           taxa2_clements %in% network_species) %>%
    left_join(interaction_categories, by = "interaction")

  #Create graph with vertex and edge attributes
  network_graph <- graph_from_data_frame(
    d = network_data[, c("taxa1_clements", "taxa2_clements")],
    vertices = NULL)

  V(network_graph)$is_focal <- V(network_graph)$name == focal_species
  E(network_graph)$interaction <- network_data$interaction
  
  #Add category attribute if using category colors
  if (cats) {
    E(network_graph)$category <- network_data$category
  }

  #Get bird information
  bird_info <- get_avicommons_image(focal_species, size = image_size)

  #Build network plot
  network_plot <- ggraph(network_graph, layout = layout)

  #Add edges - color by category or interaction type
  if (curve_edges) {
    if (cats) {
      network_plot <- network_plot +
        geom_edge_fan(aes(color = category),
                      strength = 0.75,
                      width = edge_width,
                      alpha = edge_alpha)
    } else {
      network_plot <- network_plot +
        geom_edge_fan(aes(color = interaction),
                      strength = 0.75,
                      width = edge_width,
                      alpha = edge_alpha)
    }
  } else {
    if (cats) {
      network_plot <- network_plot +
        geom_edge_link(aes(color = category),
                       width = edge_width,
                       alpha = edge_alpha)
    } else {
      network_plot <- network_plot +
        geom_edge_link(aes(color = interaction),
                       width = edge_width,
                       alpha = edge_alpha)
    }
  }

  #Apply color scale based on cats parameter
  if (cats) {
  network_plot <- network_plot +
    scale_edge_color_manual(values = setNames(interaction_categories$category_color,
                                              interaction_categories$category))
  } else {
    network_plot <- network_plot +
      scale_edge_color_manual(values = setNames(interaction_categories$color,
                                                interaction_categories$interaction))
  }

  #Add nodes
  network_plot <- network_plot +
    geom_node_point(data = function(x) x[x$is_focal, ],
                    shape = 23,
                    size = focal_node_size,
                    stroke = focal_node_stroke,
                    color = "black",
                    fill = "red") +
    geom_node_point(data = function(x) x[!x$is_focal, ],
                    shape = 21,
                    color = "black",
                    fill = "white",
                    size = node_size,
                    stroke = node_stroke)

  #Add labels if requested
  if (show_labels) {
    network_plot <- network_plot +
      geom_node_text(data = function(x) x[x$is_focal, ],
                     aes(label = name),
                     repel = TRUE,
                     size = focal_label_size,
                     fontface = "bold",
                     color = "black") +
      geom_node_text(data = function(x) x[!x$is_focal, ],
                     aes(label = name),
                     repel = TRUE,
                     size = label_size)
  }

  #Apply theme and styling
  network_plot <- network_plot +
    theme_void() +
    labs(edge_color = "Interaction Type") +
    theme(legend.position = "right",
          legend.text = element_text(size = legend_text_size),
          legend.title = element_text(size = legend_title_size),
          plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
          plot.margin = margin(15, 15, 15, 15)) +
    guides(edge_color = guide_legend(ncol = 1))

  #Create bird image panel
  bird_plot <- ggdraw() +
    draw_image(bird_info$url, x = 0, y = 0.2, width = 1, height = 0.8) +
    draw_label(bird_info$common_name,
               x = 0.5, y = 0.26, size = 22, fontface = "bold", hjust = 0.5) +
    draw_label(bird_info$scientific_name,
               x = 0.5, y = 0.22, size = 20, fontface = "italic", hjust = 0.5) +
    draw_label(paste0("Photo: ", bird_info$photographer, " | ",
                      toupper(bird_info$license)),
               x = 0.5, y = 0.18, size = 18, hjust = 0.5)

  #Combine and return
  combined_plot <- plot_grid(
    bird_plot,
    ggdraw() + draw_plot(network_plot, x = 0.02, y = 0.02,
                         width = 0.96, height = 0.96),
    ncol = 2,
    rel_widths = c(bird_width, network_width)
  )

  final_plot <- ggdraw() +
    draw_plot(combined_plot, x = 0.02, y = 0, width = 0.96, height = 1)

  return(final_plot)
}
```

The second function takes three species, and plots their networks side-by-side, with their locations on each others' networks indicated! This was used to generate the figure in the paper:

```{r}
create_combined_network <- function(data, #dataset (inter_NA_int)
                                    focal_species_vector, #vector of three species scientific names
                                    layout = "fr", #geometric layout algorithm for plot
                                    image_size = 480, #size of bird image, options are 240, 320, 480, 900
                                    edge_alpha = 0.5, #alpha value of interactions (transparency)
                                    edge_width = 1.25, #linewidth of interactions
                                    focal_node_size = 10, #size of species of interest
                                    node_size = 8,#size of interacting species
                                    node_stroke = 1.2, #node outline width
                                    focal_node_stroke = 1.5, #focal species outline width
                                    legend_text_size = 23, #interaction type text size
                                    legend_line_size = 2, #size of lines in legend
                                    curve_edges = FALSE, #set true for visualizing multiple int. between two species
                                    cats = FALSE) { #set true to color by category instead of interaction type

  #Assign styles to each focal species
  species_styles <- data.frame(
    species = focal_species_vector,
    shape = c(23, 24, 22),
    color = c("red", "blue", "gold"),
    stringsAsFactors = FALSE
  )

  #Create lists to store components
  bird_plots <- list()
  network_plots <- list()
  all_interactions <- c()  #Track all interaction types, used later in mutual legend
  all_categories <- c()  #Track all categories when cats = TRUE

  #Loop through each focal species
  for (i in seq_along(focal_species_vector)) {
    focal_species <- focal_species_vector[i]

    #Filter interactions involving focal species
    focal_interactions <- data %>%
      filter(taxa1_clements == focal_species |
               taxa2_clements == focal_species)

    #Get network species and their interactions
    network_species <- unique(c(focal_interactions$taxa1_clements,
                                focal_interactions$taxa2_clements))

    network_data <- data %>%
      filter(taxa1_clements %in% network_species,
             taxa2_clements %in% network_species) %>%
      left_join(interaction_categories, by = "interaction")

    #Track unique interactions/categories across all networks
    all_interactions <- c(all_interactions, network_data$interaction)
    if (cats) {
      all_categories <- c(all_categories, as.character(network_data$category))
    }

    #Create graph with vertex and edge attributes
    network_graph <- graph_from_data_frame(
      d = network_data[, c("taxa1_clements", "taxa2_clements")],
      vertices = NULL
    )

    #Identify which nodes are focal species
    V(network_graph)$is_current_focal <- V(network_graph)$name == focal_species
    V(network_graph)$is_other_focal <- V(network_graph)$name %in% focal_species_vector &
      V(network_graph)$name != focal_species
    V(network_graph)$is_regular <- !V(network_graph)$is_current_focal &
      !V(network_graph)$is_other_focal

    #Assign shapes and colors
    V(network_graph)$node_shape <- sapply(V(network_graph)$name, function(sp) {
      if (sp %in% species_styles$species) {
        species_styles$shape[species_styles$species == sp]
      } else {
        21 #circle for non-focal species
      }
    })

    V(network_graph)$node_color <- sapply(V(network_graph)$name, function(sp) {
      if (sp %in% species_styles$species) {
        species_styles$color[species_styles$species == sp]
      } else {
        "white" #white color for non-focal species
      }
    })

    E(network_graph)$interaction <- network_data$interaction

    #Add category attribute if using category colors
    if (cats) {
      E(network_graph)$category <- as.character(network_data$category)
    }

    #Get bird information
    bird_info <- get_avicommons_image(focal_species, size = image_size)

    #Build network plot
    network_plot <- ggraph(network_graph, layout = layout)

    #Add edges - color by category or interaction type
    if (curve_edges) {
      if (cats) {
        network_plot <- network_plot +
          geom_edge_fan(aes(color = category),
                        strength = 0.75,
                        width = edge_width,
                        alpha = edge_alpha)
      } else {
        network_plot <- network_plot +
          geom_edge_fan(aes(color = interaction),
                        strength = 0.75,
                        width = edge_width,
                        alpha = edge_alpha)
      }
    } else {
      if (cats) {
        network_plot <- network_plot +
          geom_edge_link(aes(color = category),
                         width = edge_width,
                         alpha = edge_alpha)
      } else {
        network_plot <- network_plot +
          geom_edge_link(aes(color = interaction),
                         width = edge_width,
                         alpha = edge_alpha)
      }
    }

    #Apply color scale based on cats parameter
    if (cats) {
      # Get unique category-color pairs from the actual network data
      category_colors <- network_data %>%
        filter(!is.na(category)) %>%
        mutate(category = as.character(category)) %>%
        select(category, category_color) %>%
        distinct()

      network_plot <- network_plot +
        scale_edge_color_manual(values = setNames(category_colors$category_color,
                                                  category_colors$category))
    } else {
      network_plot <- network_plot +
        scale_edge_color_manual(values = setNames(interaction_categories$color,
                                                  interaction_categories$interaction))
    }

    #Add nodes - current focal species
    network_plot <- network_plot +
      geom_node_point(data = function(x) x[x$is_current_focal, ],
                      aes(shape = node_shape, fill = node_color),
                      size = focal_node_size,
                      stroke = focal_node_stroke,
                      color = "black") +
      scale_shape_identity() +
      scale_fill_identity()

    #Add nodes - other focal species
    if (any(V(network_graph)$is_other_focal)) {
      network_plot <- network_plot +
        geom_node_point(data = function(x) x[x$is_other_focal, ],
                        aes(shape = node_shape, fill = node_color),
                        size = focal_node_size * 0.8,
                        stroke = node_stroke,
                        color = "black")
    }

    #Add nodes - regular species
    network_plot <- network_plot +
      geom_node_point(data = function(x) x[x$is_regular, ],
                      shape = 21,
                      color = "black",
                      fill = "white",
                      size = node_size,
                      stroke = node_stroke)

    #Apply theme - no legend
    network_plot <- network_plot +
      theme_void() +
      theme(legend.position = "none", #this is done because a unified legend is used instead
            plot.margin = margin(10, 10, 10, 10))

    #Get the shape and color for this focal species
    focal_shape <- species_styles$shape[species_styles$species == focal_species]
    focal_color <- species_styles$color[species_styles$species == focal_species]

    #Create a small ggplot with just the symbol
    symbol_plot <- ggplot() +
      geom_point(aes(x = 0, y = 0),
                 shape = focal_shape,
                 fill = focal_color,
                 color = "black",
                 size = 8,
                 stroke = 1.5) +
      theme_void() +
      coord_cartesian(xlim = c(-1, 1), ylim = c(-1, 1))

    #Create bird image panel with symbol
    bird_plot <- ggdraw() +
      draw_image(bird_info$url, x = 0, y = 0.15, width = 1, height = 0.85) +
      draw_plot(symbol_plot, x = 0.68, y = 0.14, width = 0.08, height = 0.16) +
      draw_label(bird_info$common_name,
                 x = 0.5, y = 0.08, size = 26, fontface = "bold", hjust = 0.5) +
      draw_label(bird_info$scientific_name,
                 x = 0.5, y = -0.02, size = 24, fontface = "italic", hjust = 0.5)

    #Store components
    bird_plots[[i]] <- bird_plot
    network_plots[[i]] <- network_plot
  }

  #Create legend based on cats parameter
  if (cats) {
    #Get unique categories present across all three networks
    unique_categories <- unique(all_categories)

    #Filter interaction_categories to only include present categories
    legend_data <- interaction_categories %>%
      select(category, category_color) %>%
      distinct() %>%
      mutate(category = as.character(category)) %>%
      filter(category %in% unique_categories)

    legend_data$category <- factor(legend_data$category,
                                   levels = legend_data$category)

    #Create a standalone legend plot with LINES
    legend_plot <- ggplot(legend_data, aes(x = category, y = 1, color = category)) +
      geom_line(size = legend_line_size) +
      scale_color_manual(values = alpha(setNames(legend_data$category_color, legend_data$category), edge_alpha),
                         name = "Interaction Category") +
      theme_void() +
      theme(legend.position = "bottom",
            legend.text = element_text(size = legend_text_size),
            legend.box.margin = margin(t = 10, b = 5),
            legend.title = element_blank()) +
      guides(color = guide_legend(nrow = 2, byrow = TRUE, override.aes = list(size = 2)))

  } else {
    #Get unique interactions present across all three networks
    unique_interactions <- unique(all_interactions)

    #Filter interaction_categories to only include present interactions
    legend_data <- interaction_categories %>%
      filter(interaction %in% unique_interactions)

    legend_data$interaction <- factor(legend_data$interaction,
                                      levels = legend_data$interaction)

    #Create a standalone legend plot with LINES
    legend_plot <- ggplot(legend_data, aes(x = interaction, y = 1, color = interaction)) +
      geom_line(size = legend_line_size) +
      scale_color_manual(values = alpha(setNames(legend_data$color, legend_data$interaction), edge_alpha),
                         name = "Interaction Type") +
      theme_void() +
      theme(legend.position = "bottom",
            legend.text = element_text(size = legend_text_size),
            legend.box.margin = margin(t = 10, b = 5),
            legend.title = element_blank()) +
      guides(color = guide_legend(nrow = 4, byrow = TRUE, override.aes = list(size = 2)))
  }

  #Extract just the legend
  legend_grob <- cowplot::get_legend(legend_plot)

  #Arrange bird images in a row
  bird_row <- plot_grid(plotlist = bird_plots, ncol = 3, rel_widths = c(1, 1, 1))

  #Arrange network plots in a row
  network_row <- plot_grid(plotlist = network_plots, ncol = 3, rel_widths = c(1, 1, 1))

  #Add legend below networks
  network_with_legend <- plot_grid(network_row, legend_grob,
                                   ncol = 1,
                                   rel_heights = c(1, 0.15))

  #Combine birds and networks vertically
  combined_plot <- plot_grid(bird_row, network_with_legend,
                             ncol = 1,
                             rel_heights = c(0.2, 0.8))

  #Add dashed vertical lines between each bird/network column
  final_plot <- ggdraw(combined_plot) +
    geom_segment(aes(x = 1/3, xend = 1/3, y = 0.15, yend = 1),
                 linetype = "dashed", color = "black", linewidth = 0.8) +
    geom_segment(aes(x = 2/3, xend = 2/3, y = 0.15, yend = 1),
                 linetype = "dashed", color = "black", linewidth = 0.8)

  return(final_plot)
}
```

Now we have everything we need to generate the plots! See `8_summary_vignette.qmd`
