---
title: "L1 post-taxonomic harmonization subset to Canada/AK/CONUS"
author: 
  - Phoebe L. Zarnetske
  - Patrick Bills
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: source
date: 10/06/2025
date-format: iso
---

## Avian Interaction Pairs Data: L1 to regional subset for Canada/Alaska/CONUS

### About

- **PROJECT:** Avian Interaction Database and MetaNetwork
- **COLLABORATORS:** Vincent Miele, Stephane Dray, Emily Parker
- **Run Date:** `r Sys.Date()`

### Data inputs:

**Avian Interaction Database L0**

Raw Data, combined (stitched) and with content
cleaning but taxonomic names as entered (older names, typos, etc)

*Set the interactions L1 database file to read from here:*


( output From `R/L1/AvianInteractionData_L1.R` )

```{r}

AvianInteractionData_L1_file <-  "AvianInteractionData_L1.csv"

```

**Main Checklists:**

Curated Species checklist: 

- Current selected species list for Canada/Alaska/CONUS =
  combined checklist based on Clements/eBird v2024, Avibase v8.17, BBS v2024:
 ( output via Working repo, from `R/L1/AvianInteractionData_specieslists_Canada_CONUS_L1.R` ) = canada.conus.splist_L1.CSV


### DATA OUTPUT:

L1 Canada, Alaska, CONUS species from L1 Avian Interaction Database 
  (`AvianInteractionData_L1.csv`); saved as:

```{r}
AvianInteractionData_CanadaAKCONUS_L1_file <- "AvianInteractionData_CanadaAKCONUS_L1.csv" 
splist_CanadaAKCONUS_L1_file <- "splist_CanadaAKCONUS_L1.csv"
today_date_string <- format(Sys.Date(),"%Y-%m-%d")

```


### Next script to run: 

- depends on project: 
  for Avian Interaction Database North America:
  for Avian MetaNetwork Analysis: https://github.com/SpaCE-Lab-MSU/avian-meta-network/tree/main/R/L2


### NOTES

This script is used to subset out only the species that occur in CONUS+Alaska+Canada.  


## Data Set-up

File paths used for this run:

```{r}
#| echo: false

# functions used by all 
source(here::here('R/L0/L0_functions.R'))
# functions for this process to reduce this script size
# and load libraries used here
source(here::here('R/L1/L1_taxonomy_functions.R'))

file_paths <- get_file_paths()

print(file_paths)
```

*Note that the Checklist folder above is in L0 which contains species lists, 
but the subset list for 2024-25 analysis is in L1 as it's built from L0 lists*



### Setup: Checklists

#### Curated List

October 2025: new curated list curated by PLZ using Clements, AVIBase, BBS etc. 
See `R/L1/AvianInteractionData_specieslists_Canada_CONUS_L1.R`

```{r}

current_species_list_filename <- "canada.conus.splist_L1.csv"
splist_filepath <- file.path(file_paths$L1,"species_checklists",current_species_list_filename )
splist <- read.csv(splist_filepath)
print(paste("Read in ", nrow(splist), "entries from checklist file", splist_filepath))

```



#### Interactions Data

To use a different or test files, edit this L0 filename:

```{r}
intxns.db.file <- file.path(file_paths$L1,AvianInteractionData_L1_file)
intxns.db <-read.csv(intxns.db.file)
print(paste("read in ", nrow(intxns.db), "records from ", intxns.db.file))
```

algorithm

@billspat this is not needed bc we already have Clements columns
<!-- - normalize -->
<!--   - make list of all clements species to use from unsplitting combo list, this is the actual clements name -->
<!--   - each of those should have a list row nmuber  -->

- get all interactions where both sp1 and sp2 on on that list
- of those interactions, convert sp1/sp2 to the sci name phoebe wants 
    

Subset by scientific_name_clements2024
```{r}
length(unique(splist$scientific_name_clements2024))
dim(splist) #783
dim(intxns.db) #31040
# Some duplicate rows; these are just the unids. so we can ignore them
duplicate_indices <- duplicated(splist$scientific_name_clements2024)
duplicate_clements <- splist[duplicate_indices, ]
duplicate_clements

intxns.db.subset <- intxns.db[intxns.db$species1_scientific %in% splist$scientific_name_clements2024 | intxns.db$species2_scientific %in% splist$scientific_name_clements2024,]

intxns.db.non.subset <- intxns.db[!(intxns.db$species1_scientific %in% splist$scientific_name_clements2024 | intxns.db$species2_scientific %in% splist$scientific_name_clements2024),]

usp1<-(unique(intxns.db.subset$species1_scientific))
usp2<-(unique(intxns.db.subset$species2_scientific))
usp1.2<-append(usp1,usp2)
length(unique(usp1.2)) #2138 unique species in the interaction data

# Merge these data with a simplified Clements/eBird BBS AOU combo list:
splist<-splist %>%
      select(scientific_name_clements2024, scientific_name_clements2024.combo,
             common_name_clements2024, common_name_clements2024.combo, 
             AOU_bbs2024, AOU_bbs2024.combo, in_bbs2024)

# Update names for merge in analysis code
splist <- splist %>%
  rename(scientific_name = scientific_name_clements2024,
         scientific_name.combo = scientific_name_clements2024.combo,
         common_name = common_name_clements2024,
         common_name.combo = common_name_clements2024.combo)

```

save the file with the file name defined in the OUTPUT section above
this has the current date embedded in it


```{r}

f <- file.path(file_paths$L1, AvianInteractionData_CanadaAKCONUS_L1_file )
write_csv(intxns.db.subset, f)
print(f)

s <- file.path(file_paths$L1, splist_CanadaAKCONUS_L1_file )
write_csv(splist, s)
print(s)

```

@billspat: Could add another merge that pulls in the splist into the intxn.db.subset to create a single dataset; need to merge by sp1 and sp2

### PLZ: END OF CODE ###






did this work???
```{r}
for(s in intxns.db.subset$species1_scientific) { if(nrow(filter(sp_lookup,species == s))!=1){print(s)}  }
```

```{r}
# I hate rowwise , loops are easier to read
intxns.db.subset$sp1 <- NA
intxns.db.subset$sp2 <- NA

for(r in 1:nrow(intxns.db.subset)){
  s1 <- intxns.db.subset[r,]$species1_scientific
  r1<- unlist(sp_lookup[sp_lookup$species == s1, "row"])
  if(length(r1)!=1) {
    # yes there are duplicate matches and I'm just picking the first one to get this done
    r1 <- r1[1]
  }
  intxns.db.subset[r,]$sp1 <- splist[r1,"scientific_name_clements2024"]
  
  s2 <- intxns.db.subset[r,]$species2_scientific
  r2<- unlist(filter(sp_lookup,species == s2)$row)
  if(length(r2)!=1) {
    r2 <- r2[1]
  }
  intxns.db.subset[r,]$sp2 <- splist[r2,"scientific_name_clements2024"]
  
}


```

Now columns sp1 and sp1 match the field scientific_name_clements2024 from CONUS etc list

# TODO #####################

$species1_scientific etc are THE ORIGINAL NAMES, NOT CLEMENTS OR BBS

select, rename etc fields on this database to match what you need here

```{r}


```







