---
title: "L1 post-taxonomic harmonization subset to Canada/AK/CONUS"
author: 
  - Phoebe L. Zarnetske
  - Patrick Bills
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor: source
date: 10/06/2025
date-format: iso
---

## Avian Interaction Pairs Data: L1 to regional subset for Canada/Alaska/CONUS

### About

- **PROJECT:** Avian Interaction Database and MetaNetwork
- **COLLABORATORS:** Vincent Miele, Stephane Dray, Emily Parker
- **Run Date:** `r Sys.Date()`

### Data inputs:

**Avian Interaction Database L0**

Raw Data, combined (stitched) and with content
cleaning but taxonomic names as entered (older names, typos, etc)

*Set the interactions L1 database file to read from here:*


( output From `R/L1/AvianInteractionData_L1.R` )

```{r}

AvianInteractionData_L1_file <-  "AvianInteractionData_L1.csv"

```

**Main Checklists:**

Curated Species checklist: 

- Current selected species list for Canada/Alaska/CONUS =
  combined checklist based on Clements/eBird v2024, Avibase v8.17, BBS v2024:
 ( output via Working repo, from `R/L1/AvianInteractionData_specieslists_Canada_CONUS_L1.R` ) = canada.conus.splist_L1.CSV


### DATA OUTPUT:

L1 Canada, Alaska, CONUS species from L1 Avian Interaction Database 
  (`AvianInteractionData_L1.csv`); saved as:

```{r}
AvianInteractionData_CanadaAKCONUS_L1_file <- "AvianInteractionData_Canada-AK-CONUS_L1.csv" 

today_date_string <- format(Sys.Date(),"%Y-%m-%d")

```


### Next script to run: 

- depends on project: 
  for Avian Interaction Database North America:
  for Avian MetaNetwork Analysis: https://github.com/SpaCE-Lab-MSU/avian-meta-network/tree/main/R/L2


### NOTES

This script is used to subset out only the species that occur in CONUS+Alaska+Canada.  


## Data Set-up

File paths used for this run:

```{r}
#| echo: false

# functions used by all 
source(here::here('R/L0/L0_functions.R'))
# functions for this process to reduce this script size
# and load libraries used here
source(here::here('R/L1/L1_taxonomy_functions.R'))

file_paths <- get_file_paths()

print(file_paths)
```

*Note that the Checklist folder above is in L0 which contains species lists, 
but the subset list for 2024-25 analysis is in L1 as it's built from L0 lists*



### Setup: Checklists

#### Curated List

October 2025: new curated list curated by PLZ using Clements, AVIBase, BBS etc. 
See `R/L1/AvianInteractionData_specieslists_Canada_CONUS_L1.R`

```{r}

current_species_list_filename <- "canada.conus.splist_L1.csv"
splist_filepath <- file.path(file_paths$L1,"species_checklists",current_species_list_filename )
splist <- read.csv(splist_filepath)
print(paste("Read in ", nrow(splist), "entries from checklist file", splist_filepath))

```



#### Interactions Data

To use a different or test files, edit this L0 filename:

```{r}
intxns.db.file <- file.path(file_paths$L1,AvianInteractionData_L1_file)
intxns.db <-read.csv(intxns.db.file)
print(paste("read in ", nrow(intxns.db), "records from ", intxns.db.file))
```

algorithm

- normalize
  - make list of all clements species to use from unsplitting combo list, this is the actual clements name
  - each of those should have a list row nmuber 

- get all interactions where both sp1 and sp2 on on that list
- of those interactions, convert sp1/sp2 to the sci name phoebe wants 
    


```{r}


splist <- dplyr::mutate(splist, ID = as.character(row_number()))
sp_lookup <- list()


for(r in 1:nrow(splist)){
  species_to_match <- str_split( splist$scientific_name[r], "; ")
  for(species in unlist(species_to_match)){
     sp_lookup<- rbind(sp_lookup, list(species, unlist(r)))
  }
}

sp_lookup<- data.frame(sp_lookup)
names(sp_lookup)<- c("species", "row")

```

subset 
```{r}
intxns.db.subset <- intxns.db[intxns.db$species1_scientific %in% sp_lookup$species & intxns.db$species2_scientific %in% sp_lookup$species,]


```

did this work???
```{r}
for(s in intxns.db.subset$species1_scientific) { if(nrow(filter(sp_lookup,species == s))!=1){print(s)}  }
```

```{r}
# I hate rowwise , loops are easier to read
intxns.db.subset$sp1 <- NA
intxns.db.subset$sp2 <- NA

for(r in 1:nrow(intxns.db.subset)){
  s1 <- intxns.db.subset[r,]$species1_scientific
  r1<- unlist(sp_lookup[sp_lookup$species == s1, "row"])
  if(length(r1)!=1) {
    # yes there are duplicate matches and I'm just picking the first one to get this done
    r1 <- r1[1]
  }
  intxns.db.subset[r,]$sp1 <- splist[r1,"scientific_name_clements2024"]
  
  s2 <- intxns.db.subset[r,]$species2_scientific
  r2<- unlist(filter(sp_lookup,species == s2)$row)
  if(length(r2)!=1) {
    r2 <- r2[1]
  }
  intxns.db.subset[r,]$sp2 <- splist[r2,"scientific_name_clements2024"]
  
}


```

Now columns sp1 and sp1 match the field scientific_name_clements2024 from CONUS etc list

# TODO #####################

$species1_scientific etc are THE ORIGINAL NAMES, NOT CLEMENTS OR BBS

select, rename etc fields on this database to match what you need here

```{r}


```


save the file with the file name defined in the OUTPUT section above
this has the current date embedded in it


```{r}

f <- file.path(file_paths$L1, paste("test_", AvianInteractionData_CanadaAKCONUS_L1_file ))
write_csv(intxns.db.subset, f)
print(f)

```







