
---
title: "Discovery Taxonomy Ammendments"
output: html_notebook
editor_options:
  chunk_output_type: inline
---

# Taxonomy Ammendments
## Avian Interaction Database

Notes and decisions made to record which species are 

See the file [taxonomy_ammendments.csv](taxonomy_ammendments.csv) for corrections
to be applied

## read in db

## blanks

display database rows with empty scientific names 
```{r}
int.raw[is.na(int.raw$species1_scientific),] # Fixed
```

## Common Names


**"Rhinoceros Auklet" should be "Cerorhinca monocerata"**

The search conditions should 
```{r}

subset(checklist, common_name == "Rhinoceros Auklet")

search_conditions <- '(species1_common == "Rhinoceros Auklet") &
(species2_common == "hybrid Western x Glaucous-winged Gull") &
(species2_scientific == "Larus occidentalis x glaucescens")
'


int.raw$species1_scientific[
  with(
    int.raw,
    (species1_common == "Rhinoceros Auklet") &
      (species2_common == "hybrid Western x Glaucous-winged Gull") &
      (species2_scientific == "Larus occidentalis x glaucescens")
  )
] <- "Cerorhinca monocerata"

