---
title: 'AvianMetaNet: Data File Repository Reporting'
author: 
  - Phoebe L. Zarnetske
  - Patrick Bills
  - Emily Parker
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
editor: source
date: 08/27/2025
date-format: iso
---

*Run date: `{r} Sys.Date()`*

# L0 Data Preparation: Stitch Raw Species Files

- PROJECT: AvianMetaNet & Avian Meta-Network
- AUTHORS: Phoebe Zarnetske, Pat Bills, Emily Parker
- COLLABORATORS: Vincent Miele, Stephane Dray
- DATA INPUT: Data entry CSV files in L0 folder
- DATA OUTPUT: L0 data: AvianInteractionData_L0.csv
- DATE: 20 Mar 2023 - AUGUST 2025
- Next script to run: 

## Report of files present in L0 folders

Data is collected in the AvianMetaNet in the form of CSV files named for the species, and checked into a git repository for tracking.  See [data entry protocol](#).  However to make it easy to enter these data, we are using Google Sheets spreadsheet for the entry, and then saving as CSVs into a "to be reviewed" folder.  CSV species data files in this folder is checked by a data manager and moved to the main 'species' folder.   

This notebook counts files in various states and checks if shows he counts of files, whichspecies and their locations.  Run this to count the current state. 

### Requirements: 

You must have access to the google drive folder, that is, the google drive software installed and connected to this computer.  You must have access to our working git repository with in-progress CSV data files, and have cloned the repository on this computer, and updated it recently.  You must configure this file or your environment to the paths where google drive files and CSV files are stored.   

```{r}
#| echo: false
#| warnings: false
source(here::here('R/lib/shared_functions.R'))

suppressMessages({
library(magrittr)
library(dplyr)
library(stringr)
})
  
# library(tidyverse)

```

## Folder path configuration

Setup the paths to the folders, requires the file 'config.yml' set in home directory and 
the code will raise warnings if it isn't.  See [readme](../../readme.md) for details. 

To use this notebook on Mac, you have to allow full access to Rstudio and/or Terminal 
 in System Settings > Privacy & Security > Full Disk Access

see https://apple.stackexchange.com/questions/439714/access-google-drive-drive-from-terminal


```{r}

#| echo: false

file_paths <- get_file_paths() 

# renaming/assigning vars for consistency with previous code

L0_dir <- file_paths$L0
species_dir<- file.path(file_paths$L0, "species")
git_repo_folder <- species_dir
species_temp_dir<-file.path(file_paths$L0, "species_temp")
species_in_review_dir<- file.path(file_paths$L0, "species_in_review")


```

File Paths used in this run: 

```{r} 
file_paths
```



*Note 12/9/2024: some EP gsheets still need checking but contain useful data. We
don't want to miss any in /species/ that are only in gsheet form.*

```{r}

# List all filenames in both folders

# this requires the google drive folder to be set
if("SYNC_DATA_FOLDER" %in% names(file_paths)){
  google_drive_folder <- file.path(file_paths$SYNC_DATA_FOLDER, "species_completed")
  google_drive_files <- list.files(google_drive_folder, full.names = FALSE)
  google_drive_main_names <- sapply(google_drive_files, get_main_name)
  google_only_main_names <- setdiff(google_drive_main_names, git_repo_main_names)
  git_only_main_names <- setdiff(git_repo_main_names, google_drive_main_names)
  # Find common names (intersection) and unique names (differences)
  common_main_names <- intersect(google_drive_main_names, git_repo_main_names)

  print("Number of Google Only Main Names:")  
  print( length(google_only_main_names))

  tibble(google_only_main_names)

  print(paste("Git only", length(git_only_main_names),  "files "))

  tibble(git_only_main_names)

  print("Files with matching main names in both Google Drive and Git repository:")
  print(common_main_names)
}
```

In Winter 2025, those only in G-drive appear to be files that do not have initials for entry and review in the suffix, and maybe files that pre-dated using this google drive folder.   


```{r}
git_repo_files <- list.files(git_repo_folder, full.names = FALSE)
species_in_review_files <- list.files(species_in_review_dir, full.names = FALSE)

# Extract main names from filenames
git_repo_main_names <- sapply(git_repo_files, get_main_name)
species_to_review_main_names <- sapply(species_in_review_files, get_main_name)

```



### Species left to review

```{r, echo=FALSE}
print(paste("There are", length(species_to_review_main_names), "files to be reviewed"))
```

Stitch and count the unique common names within files to review

```{r}
#| echo: false

csv_file_group_name='species_in_review'
print(csv_file_group_name)

csv_file_path = file.path(file_paths$L0, "species_in_review")
# filter or add in google drive files here
in_review_file_list <- list.files(path = csv_file_path , pattern = ".*\\.csv", 
  full.names = TRUE)

paste(csv_file_path, ":", csv_file_group_name, "files to process", length(in_review_file_list))

```

## Next Steps

Open and Run/Render Notebooks in `R/L0/`





