---
title: "L0_stitching_debug"
output:  notebook
---

## Avian Meta Network: debug file read

This is to run the error correction on single files to debug either the code, 
or the data, or both

### Setup

```{r}
#| echo: false
#| warnings: false

source(here::here('R/lib/shared_functions.R'))

# This uses functions in R/lib/shared_functions.R which also imports R/lib/config.R
# which sets the locations of the data folders. 

# this uses the files paths found in /filepaths.R
# edit the code above, or specify a different configuration file to use a 
# different location. 

file_paths <- get_file_paths() 
file_paths

# or override file paths here if desired
```

### Set the file to check here manually. 

**Include the species folder!**  But not the L0 folder (set above). For example, 

 > `species/lanius_borealis.csv`
 
This will allow you to run this on files to be checked

```{r}
L0_file <- "species/Calidris_ptilocnemis_EP_IH.csv"
```

Add path and check
```{r}
#| echo: false

L0_file_path <- file.path(file_paths$L0, L0_file)
print(L0_file_path)
stopifnot(file.exists(L0_file_path))

```

### Test CSV read

Examine the tibble for data in folders

```{r}
#| echo: false
test.df <- read_csv(L0_file_path)
tibble(test.df)
```


### Test Avian Meta Network Read 
```{r}

amn.df <- read_and_amend(L0_file_path,add_entry_file_column= TRUE, fix_errors_on_read = TRUE)
print(names(amn.df))
```

```{r}
#| echo: false
tibble(amn.df)
```


test the 'stitch' on this one file.  this does essentially what the process above
does, but combines files.  Since this is one file, should be the same. 

```{r}
intxnsL0sp <- L0_stitch(
                csv_file_list = list(L0_file_path), 
                csv_file_group_name = "debug",
                add_entry_file_column = TRUE, 
                fix_errors_on_read = TRUE
                )

nrow(intxnsL0sp$intxns)
```

debug wide-to-long conversion

```{r}
long.df <- sources_wide_to_long(intxnsL0sp$intxns)
tibble(long.df)
# tibble(intxnsL0sp$intxns)
```
```{r}
nrow(dplyr::filter(long.df, is.na(source_URL)))
```

```{r}

View(long.df)
```
